Echoes from the scattering of wavepackets on wormholes

### 2018-12-13 paper.tex
no message

The long absence of observational evidence confirming the dynamical properties of spacetime has motivated a plethora of conjectures about the behaviour of gravity within and beyond \citep{Clifton:2011jh, Taliotis:2012sx, Joyce:2014kja} classical General Relativity (GR). Latterly, the potential existence of exotic compact objects (ECOs) sourced by quantum effects on gravity \citep{PhysRevLett.61.1446, Almheiri:2012rt, Mazur:2001fv} (such as wormholes, firewalls and gravastars) has captured the attention of many recent efforts \citep{Cardoso:2016oxy, Abedi:2016hgu, Abedi:2018npz}, wherein the claim is that the detection of a train of ``echoes'' isolated from the main transient of GW and with generically large amplitudes would be a clear evidence of ECOs. 
\bibliography{bibliography.bib}

>>>>

The long absence of observational evidence confirming the dynamical properties of spacetime has motivated a plethora of conjectures about the behaviour of gravity within and beyond \citep{Clifton:2011jh, Taliotis:2012sx, Joyce:2014kja} classical General Relativity (GR). Latterly, the potential existence of exotic compact objects (ECOs) sourced by quantum effects on gravity \citep{PhysRevLett.61.1446, Almheiri:2012rt, Mazur:2001fv} (such as wormholes, firewalls and gravastars) has captured the attention of many recent efforts \citep{Cardoso:2016oxy, Abedi:2016hgu, Abedi:2018npz}, wherein the principal claim is that the detection of a train of ``echoes'' isolated from the main transient of GW and with generically large amplitudes would be a clear evidence of ECOs. It is, therefore, necessary to understand (i) the mechanisms behind the production of echoes and (ii) the intensity and spectrum of the outgoing wavelets compared to the GW transient in the simplest possible setup. In this paper, we explore the generation of echoes by colliding wavepackets of scalar and tensor radiation against a traversable spherically symmetric wormhole \citep{Visser:1989kh}; which behaves just like a Fabry-Perot cavity, shares common properties with the effective potential cavities made by other ECOs, like gravastars and firewalls, and the main features of the dispersed pulses are similar to the ringdown signals after the coalescence of ECOs.

Here we consider a wormhole configuration made by the junction of two Schwarzschild geometries of identical mass at $r_0>2M$, just as shown in \citep{Cardoso:2016oxy}. In this case, the symmetry of the centrifugal barriers at $r=3M$ on each side of the throat allows us to find the reflection and transmission coefficients of the cavity. Hence, it is possible to reconstruct the spectral shape of the outgoing pulse using the geometrical optics approximation. Nevertheless, this approximation predicts an exponential decay of the subsequent higher order reflections, which appears instead as a power law in the full solution of the scattering problem. Thus, the time evolving profile of the echoes can only be generated by the excitation of quasinormal modes (QNMs), these modes are sourced by a sequence of internal reflections inside the potential cavity and then propagate throughout the surface of the maximal potential energy spheres (i.e. the ``edges'' of the potential barriers), while leaking energy into the exterior. Quasinormal modes of the Schwarzschild solution have been extensively studied and reproduced in various analytic and numerical simulations \citep{Chandrasekhar:1975zza, PhysRevD.46.4179}, thus it is easy to identify their characteristic frequencies in the spectrum of outgoing pulses. We as well present the full scattering solution both inside and outside the wormhole cavity in detail, along with the energy fluxes and the asymptotic solutions for the principal spherical modes of a scalar (and tensor) wavepackets. In addition to this, we find the width and frequency intervals contained in the incident wavepackets for which the outgoing the wavelets have maximal amplitudes.

The layout for this paper is as follows: 
\bibliography{bibliography.bib}
